---
title: "Analyses for event segmentation thesis"
author: "Jenni Saaristo"
date: "18.8.2021"
output:
  pdf_document:
    fig_crop: false
---

These are the final analyses reported in the thesis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment="")
outdir = '/Users/jenska/code/python/eventcode/5_analyses_plots/out/'
figdir <- '/Users/jenska/code/python/eventcode/5_analyses_plots/figs/' 
indir <- '/Users/jenska/code/python/eventcode/4_get_averages/out/'
suppressMessages(library(tidyverse))
```

# Main analyses

The analyses for the main research questions: responses for bounds versus perms, modulation by salience, and plotting these with the FIR models.

## Calculate significance of bounds vs perms and plot results

Define which data to use for all results:
```{r define_strategy}
re_time <- '3s'
if (re_time == '3s') {is_3s <- TRUE} else {is_3s <- FALSE}
```

### Statistical analyses
```{r boundsperms_stats}
datapath <- paste(indir,'AVG_boundperms_',re_time,'_all.csv', sep="")
data <- read_csv(datapath, col_types = cols())
areas <- unique(data$area)

results <- data.frame(area=areas, beta=0, p=1, perm_mean=0, perm_sd=0)
for (a in areas) {
  perm_betas <- data$beta[data$perm != 0 & data$area == a]
  bound_beta <- data$beta[data$perm == 0 & data$area == a]
  p <- length(perm_betas[perm_betas > bound_beta])/length(perm_betas) # in fact we just want 1-tailed
  results$beta[results$area == a] <- bound_beta
  results$p[results$area == a] <- p
  results$perm_mean[results$area == a] <- mean(perm_betas)
  results$perm_sd[results$area == a] <- sd(perm_betas)
}
results$p_adj <- p.adjust(results$p, method='holm')
results[results$area == areas[49],]

topareas <- results[results$p_adj < 0.05,]
topareas <- topareas[order(topareas$beta, decreasing = TRUE),]
topareas

for (i in 1:nrow(topareas)) {
  writeLines(paste(topareas$area[i], round(topareas$beta[i],3), round(topareas$p_adj[i],6), sep = '\t'))
}

write_csv(results, paste(outdir,'boundperm_results_',re_time,'.csv', sep=""))
sigs <- as.numeric(rownames(topareas))
sigs <- append(sigs, 49)
```

### Plot bounds vs perms
We'll only plot this for HC, the rest will make do with FIRs and stats
```{r boundsperms_plots}
# Hippocampus
a = c(49)
asel <- areas[a]
bounds <- subset(data, perm == 0 & area %in% asel)
perms <- subset(data, perm != 0 & area %in% asel)

png(paste(figdir,'bounds_vs_perms_',re_time,'_HC.png',sep=""), width=550, height=400)
ggplot(data=perms, aes(group=area)) +
  geom_histogram(aes(x=beta), binwidth = 2) +
  geom_vline(mapping=aes(xintercept=beta), data=bounds, color='red', size=2) +
  facet_wrap( ~ area,ncol=2) +
  labs(x='beta values', y='count (total 1000)', title=paste('Latency -', re_time, sep="")) +
  scale_x_continuous(limits=c(-17,17), breaks=seq(-15,15,5)) +
  theme_grey(base_size = 25)
dev.off()
```

### Plot FIRs
We'll only use the ones from 3s, with an added line for the 1s bound. In addition to HC we'll plot all areas that are significant in either condition, because the reader should have all that info available -- it sucks when you don't have it. For this first part we plot the grand average over all saliences as that's more clearly interpretable in conjunction with the cortical plots.

```{r fir_plots, eval=is_3s, include=is_3s}
datapath <- paste(indir,'AVG_fir_',re_time,'_all.csv', sep="")
data <- read_csv(datapath, col_types = cols())
data$cond <- as.factor(data$cond)
data$cond <- ordered(data$cond, levels=c('high','mid','low','all'))

data <- filter(data, grepl('delay', regressor))
data$delay <- as.numeric(gsub(data$regressor, pattern="[^0-9]", replacement=""))
data$delay <- data$delay - 8  # adjust latencies so that behav boundary is 0
areas <- unique(data$area)

# Plot HC large
conds <- c('high','mid','low')
asel <- areas[49]
plotdata <- subset(data, area %in% asel & cond %in% conds)
png(paste(figdir,'bounds_FIRs_',re_time,'_HC.png', sep=""), width=1100, height=700)
pd <- position_dodge(0.2)
ggplot(plotdata, aes(x=delay, y=beta, color=cond)) +
  geom_hline(yintercept = 0, color='grey50', size=1) +
  geom_vline(xintercept = -3, color='blue3', size=2) +   # latency = -3
  geom_vline(xintercept = -1, color='green3', size=2) +  # latency = -1
  geom_point(position=pd, size=8) +
  geom_line(position=pd, size=3) +
  geom_errorbar(aes(ymin=beta-se, ymax=beta+se), width=0.5, size=1, position=pd) +
  scale_color_grey(name='salience') +
  scale_x_continuous(breaks=seq(-8,7)) +
  labs(title=paste('Modulation by salience')) +
  xlab('delay (sec)') +
  facet_wrap(~ area, ncol=1) +
  theme_bw(base_size = 24) +
  theme(panel.grid=element_blank())
dev.off()

# Plot others small, grand average
conds <- c('all')
asel <- areas[sigs]
plotdata <- subset(data, area %in% asel & cond %in% conds)
pd <- position_dodge(0.2)
png(paste(figdir,'bounds_FIRs_',re_time, '_avg_cortical.png', sep=""), width=1400, height=700)
ggplot(plotdata, aes(x=delay, y=beta)) +
  geom_hline(yintercept = 0, color='grey50', size=1) +
  geom_vline(xintercept = -3, color='blue3', size=2) +   # latency = -3
  geom_vline(xintercept = -1, color='green3', size=2) +  # latency = -1
  geom_point(position=pd, size=3) +
  geom_line(position=pd, size=1) +
  geom_errorbar(aes(ymin=beta-se, ymax=beta+se), width=0.3, size=0.7, position=pd) +
  #scale_color_grey(name='salience') +
  scale_x_continuous(breaks=seq(-8,7)) +
  scale_y_continuous(limits=c(-7,12), breaks=seq(-5,10,5)) +
  labs(title=paste('FIR responses')) +
  xlab('delay (sec)') +
  facet_wrap(~ area, ncol=4, scales='free') +
  theme_bw(base_size = 20) +
  theme(panel.grid=element_blank())
dev.off()

```

```{r include=FALSE}
rm(data,datapath,bounds,perms,bound_beta,perm_betas,results,topareas,conds,plotdata,areas,asel,a,pd,p,plotlist,i,sigs)
```


## Modulation by salience
```{r modulation_stats}
suppressMessages(library(lme4))
suppressMessages(library(lmerTest))

datapath <- paste(indir,'betas_series_long_',re_time,'.csv', sep="")
data <- read_csv(datapath, col_types = cols())
boundpath <- paste('/Users/jenska/code/python/eventcode/1_create_boundaries/out/boundaries_f',re_time,'_vol.csv', sep="")
bounds <- read_csv(boundpath, col_types = cols())
areas <- unique(data$area)

# Salience for bounds (bins: 5-6, 7-9, 10-17)
bounds$salience <- 0
bounds$salience[bounds$nobs >= 7] <- 1
bounds$salience[bounds$nobs >= 10] <- 2

# Clean some unnecessary columns
bounds <- subset(bounds, select=c(id,nobs,salience,meanvol,voldiff))
data <- subset(data, select=-c(X1))
data <- dplyr::rename(data, id=bound)

# Also remove the post-hoc bounds from data
data <- data[data$id != 999,]

# Join the data frames
data <- inner_join(data,bounds, by="id")

# Check correlations
cor.test(bounds$salience, bounds$meanvol)
cor.test(bounds$salience, bounds$voldiff)
cor.test(bounds$nobs, bounds$meanvol)
cor.test(bounds$nobs, bounds$voldiff)

# Run glm on HC, salience
datasel <- data[data$area == areas[49],]
m1 <- lmer(beta ~ salience + meanvol + voldiff + (1 | subj) + (1 | id), datasel)
anova(m1)
m11 <- lmer(beta ~ salience + (1 | subj) + (1 | id), datasel)
anova(m11)

# Run glm on HC, nObs
datasel <- data[data$area == areas[49],]
m2 <- lmer(beta ~ nobs + meanvol + voldiff + (1 | subj) + (1 | id), datasel)
anova(m2)
m21 <- lmer(beta ~ nobs + (1 | subj) + (1 | id), datasel)
anova(m21)

# Run for all ROIs
results <- data.frame(area=areas, f_sal=0, p_sal=1, f_obs=0, p_obs=1)

for (i in c(1:49)) {
  datasel <- data[data$area == areas[i],]
  anv1 <- anova(lmer(beta ~ salience +  meanvol + voldiff + (1 | subj) + (1 | id), datasel))
  anv2 <- anova(lmer(beta ~ nobs +  meanvol + voldiff + (1 | subj) + (1 | id), datasel))
  results$f_sal[i] <- anv1$`F value`[1]
  results$p_sal[i] <- anv1$`Pr(>F)`[1]
  results$f_obs[i] <- anv2$`F value`[1]
  results$p_obs[i] <- anv2$`Pr(>F)`[1]
}
results$p_sal_adj <- p.adjust(results$p_sal, method='holm')
results$p_obs_adj <- p.adjust(results$p_obs, method='holm')

# Select results
topareas <- results[results$p_sal_adj < 0.05 | results$p_obs_adj < 0.05,]
topareas <- topareas[order(topareas$f_obs, decreasing = T),]
topareas

results[results$area == areas[49],]

for (i in 1:nrow(topareas)) {
  writeLines(paste(topareas$area[i], round(topareas$f_sal[i],3), round(topareas$p_sal[i],3), round(topareas$p_sal_adj[i],3), round(topareas$f_obs[i],3), round(topareas$p_obs[i],3), round(topareas$p_obs_adj[i],3), sep = '\t'))
}
sigs <- as.numeric(rownames(topareas))
sigs <- append(sigs, 49)
```

Without confounds
```{r modulation_stats_nocon}

# Run for all ROIs
results_nocon <- data.frame(area=areas, f_sal=0, p_sal=1, f_obs=0, p_obs=1)

for (i in c(1:49)) {
  datasel <- data[data$area == areas[i],]
  anv1 <- anova(lmer(beta ~ salience + (1 | subj) + (1 | id), datasel))
  anv2 <- anova(lmer(beta ~ nobs + (1 | subj) + (1 | id), datasel))
  results_nocon$f_sal[i] <- anv1$`F value`[1]
  results_nocon$p_sal[i] <- anv1$`Pr(>F)`[1]
  results_nocon$f_obs[i] <- anv2$`F value`[1]
  results_nocon$p_obs[i] <- anv2$`Pr(>F)`[1]
}
results_nocon$p_sal_adj <- p.adjust(results_nocon$p_sal, method='holm')
results_nocon$p_obs_adj <- p.adjust(results_nocon$p_obs, method='holm')

# Select results
topareas_nocon <- results_nocon[results_nocon$p_sal_adj < 0.05 | results_nocon$p_obs_adj < 0.05,]
topareas_nocon <- topareas_nocon[order(topareas_nocon$f_obs, decreasing = T),]
topareas_nocon

results_nocon[results_nocon$area == areas[49],]

for (i in 1:nrow(topareas_nocon)) {
  writeLines(paste(topareas_nocon$area[i], round(topareas_nocon$f_sal[i],3), round(topareas_nocon$p_sal[i],3), round(topareas_nocon$p_sal_adj[i],3), round(topareas_nocon$f_obs[i],3), round(topareas_nocon$p_obs[i],3), round(topareas_nocon$p_obs_adj[i],3), sep = '\t'))
}
sigs_nocon <- as.numeric(rownames(topareas_nocon))
sigs_nocon <- append(sigs, 49)

```



### Plot FIRs
With salience
```{r, eval=is_3s, include=is_3s}
datapath <- paste(indir,'AVG_fir_',re_time,'_all.csv', sep="")
data <- read_csv(datapath, col_types = cols())
data$cond <- as.factor(data$cond)
data$cond <- ordered(data$cond, levels=c('high','mid','low','all'))

data <- filter(data, grepl('delay', regressor))
data$delay <- as.numeric(gsub(data$regressor, pattern="[^0-9]", replacement=""))
data$delay <- data$delay - 8  # adjust latencies so that behav boundary is 0
areas <- unique(data$area)
conds <- c('high','mid','low')

# Plot all with salience
asel <- areas[sigs]
plotdata <- subset(data, area %in% asel & cond %in% conds)
pd <- position_dodge(0.2)
png(paste(figdir,'bounds_FIRs_',re_time, '_sal_cortical.png', sep=""), width=1400, height=500)
ggplot(plotdata, aes(x=delay, y=beta, color=cond)) +
  geom_hline(yintercept = 0, color='grey50', size=1) +
  geom_vline(xintercept = -3, color='blue3', size=2) +   # latency = -3
  #geom_vline(xintercept = -1, color='green3', size=2) +  # latency = -1
  geom_point(position=pd, size=3) +
  geom_line(position=pd, size=1) +
  geom_errorbar(aes(ymin=beta-se, ymax=beta+se), width=0.3, size=0.7, position=pd) +
  scale_color_grey(name='salience') +
  scale_x_continuous(breaks=seq(-8,7)) +
  scale_y_continuous(limits=c(-7,18), breaks=seq(-5,15,5)) +
  labs(title=paste('FIR responses')) +
  xlab('delay (sec)') +
  facet_wrap(~ area, ncol=4, scales='free') +
  theme_bw(base_size = 20) +
  theme(panel.grid=element_blank())
dev.off()
```

```{r include=FALSE}
rm(bounds,data,datasel,areas,results,m1,m11,m2,m21,anv1,anv2,topareas,i,boundpath,datapath,plotdata,pd,conds,sigs,asel)
```


# Secondary analyses: familiarity effects
Nothing is even close to significant at 3s, even less at 1s -- does it make sense to even plot any of these? Well maybe HC has theoretical interest, so we should mention it.
```{r familiarity}
library(effsize)
datapath <- paste(indir,'betas_boundperms_long_',re_time,'.csv', sep="")
data <- read_csv(datapath, col_types = cols())
areas <- unique(data$area)

# Get subj info and merge
subjdata <- read_csv('code/python/eventcode/subj_info.csv', col_types = cols())
data <- merge(data, subjdata)

results <- data.frame(area = areas, mean_lis1=0, mean_lis2=0, t=0, p=1, d=0)

for (a in areas) {
  lis1_betas <- data$beta[data$listening == 'first' & data$area == a]
  lis2_betas <- data$beta[data$listening == 'second' & data$area == a]
  stat <- t.test(lis1_betas,lis2_betas)
  eff <- cohen.d(lis1_betas, lis2_betas)
  results$mean_lis1[results$area == a] <- stat$estimate[1]
  results$mean_lis2[results$area == a] <- stat$estimate[2]
  results$t[results$area == a] <- stat$statistic
  results$p[results$area == a] <- stat$p.value
  results$d[results$area == a] <- eff$estimate
}
results$p_adj <- p.adjust(results$p, method='holm')

rm(lis1_betas,lis2_betas,stat,eff)

results[results$area == areas[49],]

topareas <- results[results$d > 0.3,]
topareas <- topareas[order(topareas$p, decreasing = FALSE),]
topareas

for (i in 1:nrow(topareas)) {
  writeLines(paste(topareas$area[i], round(topareas$mean_lis1[i],3), round(topareas$mean_lis2[i],3),  round(topareas$d[i],3), round(topareas$p[i],3), round(topareas$p_adj[i],3), sep = '\t'))
}
```

## Plot swarms
I.e. the values that went into the t-test
```{r, eval=is_3s, include=is_3s}
library(ggbeeswarm)

asel <- areas[49]
plotdata <- subset(data, area %in% asel)
png(paste(figdir,'fam_swarm_',re_time, '_HC.png', sep=""), width=600, height=500)
ggplot(plotdata, aes(x=listening, y=beta)) +
  geom_violin(color='grey40', fill='bisque2', size=1, draw_quantiles=0.5) +
  geom_quasirandom(dodge.width=1, size=4) +
  labs(title=paste('Effect of familiarity on HC response')) +
  xlab('Listening') +
  theme_bw(base_size = 24) +
  theme(panel.grid=element_blank())
dev.off()
```

```{r include=FALSE}
rm(data,results,subjdata,topareas,a,areas,i)
```

## Plot FIRs by grp
For HC
```{r firs_fam, eval=is_3s, include=is_3s}
datapath1 <- paste(indir,'AVG_fir_',re_time,'_1st.csv', sep="")
data1 <- read_csv(datapath1, col_types = cols())
datapath2 <- paste(indir,'AVG_fir_',re_time,'_2nd.csv', sep="")
data2 <- read_csv(datapath2, col_types = cols())

data1 <- subset(data1, cond=='all')
data2 <- subset(data2, cond=='all')
data1$lis <- '1st'
data2$lis <- '2nd'
data <- rbind(data1, data2)
rm(datapath1,datapath2,data1,data2)

data <- filter(data, grepl('delay', regressor))
data$delay <- as.numeric(gsub(data$regressor, pattern="[^0-9]", replacement=""))
data$delay <- data$delay - 5

areas <- unique(data$area)

# Plot HC
asel <- areas[49]
plotdata <- subset(data, area %in% asel)
png(paste(figdir,'fam_FIRs_',re_time,'_HC.png', sep=""), width=1000, height=600)
pd <- position_dodge(0.2)
ggplot(plotdata, aes(x=delay, y=beta, color=lis)) +
  geom_hline(yintercept = 0, color='grey50', size=1) +
  geom_vline(xintercept = 0, color='blue3', size=2) +  # delay = 0
  geom_point(position=pd, size=8) +
  geom_line(position=pd, size=3) +
  geom_errorbar(aes(ymin=beta-se, ymax=beta+se), width=0.5, size=1, position=pd) +
  scale_color_grey(name='Listening') +
  scale_x_continuous(breaks=seq(-5,10)) +
  labs(title=paste('Modulation by familiarity')) +
  xlab('delay (sec)') +
  #facet_wrap(~ area, ncol=1) +
  theme_bw(base_size = 24) +
  theme(panel.grid=element_blank())
dev.off()
```

```{r include=FALSE}
rm(data,plotdata,pd,asel,areas)
```


# Test bounds against non-bounds (audio gaps)
```{r test_audiobounds, eval=is_3s, include=is_3s}
datapath <- paste(indir,'AVG_audioperms_',re_time,'_all.csv', sep="")
data <- read_csv(datapath, col_types = cols())
areas <- unique(data$area)

results <- data.frame(area=areas, beta=0, p=1, perm_mean=0, perm_sd=0)
for (a in areas) {
  perm_betas <- data$beta[data$perm != 0 & data$area == a]
  bound_beta <- data$beta[data$perm == 0 & data$area == a]
  p <- length(perm_betas[perm_betas > bound_beta])/length(perm_betas) # in fact we just want 1-tailed
  results$beta[results$area == a] <- bound_beta
  results$p[results$area == a] <- p
  results$perm_mean[results$area == a] <- mean(perm_betas)
  results$perm_sd[results$area == a] <- sd(perm_betas)
}
results$p_adj <- p.adjust(results$p, method='holm')
results[results$area == areas[49],]

topareas <- results[results$p_adj < 0.05,]
topareas <- topareas[order(topareas$beta, decreasing = TRUE),]
topareas

for (i in 1:nrow(topareas)) {
  writeLines(paste(topareas$area[i], round(topareas$beta[i],3), round(topareas$p_adj[i],6), sep = '\t'))
}

write_csv(results, paste(outdir,'audioperm_results_',re_time,'.csv', sep=""))
sigs <- as.numeric(rownames(topareas))
```

### Plot
Not sure if it makes sense to plot these... well maybe yes, there's room in the fig
```{r audioperms_plots, eval=is_3s, include=is_3s}
# Plot HC
a = c(49)
asel <- areas[a]
bounds <- subset(data, perm == 0 & area %in% asel)
perms <- subset(data, perm != 0 & area %in% asel)
png(paste(figdir,'bounds_vs_audio_',re_time,'_HC.png',sep=""), width=500, height=400)
ggplot(data=perms, aes(group=area)) +
  geom_histogram(aes(x=beta), binwidth = 2) +
  geom_vline(mapping=aes(xintercept=beta), data=bounds, color='red', size=2) +
  facet_wrap( ~ area,ncol=2) +
  labs(x='beta values', y='count (total 1000)', title='Boundaries vs. audiogaps') +
  theme_grey(base_size = 25)
dev.off()

# Plot others
asel <- areas[sigs]
bounds <- subset(data, perm == 0 & area %in% asel)
perms <- subset(data, perm != 0 & area %in% asel)
png(paste(figdir,'bounds_vs_audio_',re_time, '_cortical.png', sep=""), width=1300, height=400)
ggplot(data=perms, aes(group=area)) +
  geom_histogram(aes(x=beta), binwidth = 2) +
  geom_vline(mapping=aes(xintercept=beta), data=bounds, color='red', size=1.5) +
  facet_wrap( ~ area,nrow=2) +
  labs(x='beta values', y='count (total 1000)') +
  theme_grey(base_size = 19)
dev.off()

```

```{r include=FALSE, eval=is_3s, include=is_3s}
rm(data,bounds,perms,results,topareas,areas,a,asel,bound_beta,perm_betas,i,datapath,plotlist,p,sigs)
```